Delay d'ETRE: Explict Temporal Ratios Lead to Impatience

### 2015-12-15 paper/paperintro.tex
now with study 1 almost done

%opening
\title{To Ship or Not to Ship: Explict Temporal Ratios Lead to Impatience}
\author{Daniel Wall}

@@

How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
%For instance \citeA{Read2013} found that by making ratios in the 
However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

@@

Intertemporal preferences measured in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
There are four general categories of research on  intertemporal choices (see \citeA{Read2013} for an overview).
The first focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
 the current paper belongs to research investigating how patience is altered by option frames. 
 The second stream focuses on how choices relate to real world behavior.
 Individual differences in imputed discount functions predict real world phenomenon such as credit card debt,  mortgage choice, and smoking \cite{MacKillop2011, Meier2009, Johnson2011}. 
 The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being seeing freshly baked bread \cite{Li2008}. 
 The fourth stream investigates how the framing of options affects patience.

@@

\subsubsection{Hypothesis 1}
According to the ITCH model people rely on ratios and differences in temporal choices. 
A prediction of the model is that the higher the weight given to the temporal ratio the more impatient participants will be.

@@

\subsubsection{Hypothesis 3}
Similar to hypothesis 2 when framed as a shipping scenario making temporal ratios explicit people will be more impatient.

@@

I test these hypotheses across three studies.
Study 1 tests Hypothesis 1 by making temporal ratios explicit in intertemporal choices.

Study 2 extends the results of Study 1 to losses. 
Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
Study 2 investigates how framing times as ratios leads to decreased patience in losses.

Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
Shipping decisions can be thought of as paying more now to receive an item sooner whereas the intertemporal choices shown in Study 2 can be thought of as paying something smaller now to avoid paying something larger later.

\section{Study 1}
In study 1 participants chose between smaller sooner and larger later options. 
These options were displayed in one of two frames: implicit temporal ratios and explicit temporal ratios. 
Study 1 tests Hypothesis 1: In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.

\subsection{Method}
Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
Further, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
Additionally only participants who passed an attention check were included in the final analysis.
These precautions were taken to ensure that participants completed

Participants completed informed consent, then were given a test to ensure they understood the paradigm.
They then were assigned to one of the two conditions: implicit ratio or explicit ratio. 
They then were shown the intertemporal choices in Table []. 
Each intertemporal choice was shown in the following format.

@@

Insert a Qualtrics picture here.

Choices were displayed in a random order to each participant.

After completing their choices participants completed demographic information.

@@

The dependent variable in this experiment was the number of times a participant chose the larger later -- e.g. more patient -- reward.

each participant chose the larger-later reward ?LL?.As Table 2 shows, there was a strong date/delay effect: Twice as many people chose LL in the Date condition than in the Week or Month conditions. Differences across conditions were tested formally with Kruskal- Wallis tests. An overall test confirmed a signifi- cant difference between the conditions (?2?2? = 21?1,

@@

\begin{table}[!htbp] \centering 
	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
\begin{tabular}{ | l | l | l | l | l | l | l | l | }
	\hline
	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
\end{tabular}
\end{table}

>>>>

%opening
\title{Delay d'ETRE: Explict Temporal Ratios Lead to Impatience}
\author{Daniel Wall}

@@

How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
%For instance \citeA{Read2013} found that by making ratios in the 
However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
Further although shipping scenarios are a form of intertemporal choice, they differ from how most studies operationalize intertemporal choices. 
Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
As opposed to other studies which showed that reframed times as dates increases patience, framing times as ratios decreases patience.
Firms wanting to decrease patience in their consumers may want to frame intertemporal choices as ratios as opposed to delays.

%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

@@

Intertemporal preferences measured in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
There are four general categories of research on  intertemporal choices (see \citeA{Read2013} for an overview).
The first focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
the current paper belongs to research investigating how patience is altered by option frames. 
The second stream focuses on how choices relate to real world behavior.
Individual differences in imputed discount functions predict real world phenomenon such as credit card debt,  mortgage choice, and smoking \cite{MacKillop2011, Meier2009, Johnson2011}. 
The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being shown freshly baked bread \cite{Li2008}. 
The fourth stream investigates how the framing of options affects patience.

@@

\subsubsection{Hypothesis 1}
Given that people rely on differences and ratios between amounts and times in intertemporal choice, increasing the reliance on ratios for times with relatively small differences should lead to decreased patience.

@@

\subsubsection{Hypothesis 3}
Related to hypothesis 2 when framed an intertemporal chioce is framed as a shipping scenario,  making temporal ratios explicit people will induce less patience.

@@

\subsection{Summary of Studies}

In this paper, I describe three studies investigating the explicit temporal ratio effect (ETRE).
Study 1 shows that the explicit temporal ratios lead to less patience when intertemporal choices are framed as gains.
Study 2 extends this effect to losses.
Finally, Study 3 shows that in an explicit shipping scenario people are less patient when temporal ratios are made explicit.

%Study 2 extends the results of Study 1 to losses. 
%Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
%For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
%Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
%Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
%Study 2 investigates how framing times as ratios leads to decreased patience in losses. 
%
%Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
%Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
%Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
%Shipping decisions can be thought of as \textit{paying more now to receive an item sooner} whereas the intertemporal choices shown in Study 2 can be thought of as \textit{paying something smaller now to avoid paying something larger later}.
%This difference in choices  
%

\section{Study 1}
Study 1 shows that the explicit temporal ratios lead to less patience when intertemporal choices are framed as gains.
Participants chose between smaller sooner and larger later options. 
These options were displayed in one of two frames: implicit temporal ratios and explicit temporal ratios. 
Study 1 tests Hypothesis 1: In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.

\subsection{Method}
Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
Further, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
Additionally only participants who passed an attention check were included in the final analysis.
These precautions were taken to ensure that participants were paying attention to the study

Participants completed informed consent, then were given a test to ensure they understood the paradigm.
Next they were assigned to one of the two conditions: implicit ratio or explicit ratio. 
They then were shown the five intertemporal choices in from the Study 1 section of  Table~\ref{tab:stimuli}. 
Choices were displayed in a randomized order

@@

For the implicit time ratio condition, intertemporal choice A was displayed as follows:

\includegraphics[]{study1_implicit}

Choices B-E were displayed in the identical format.

In the explicit time ratio condition, intertemporal choice A was displayed as follows:

\includegraphics[]{study1_explicit}

Choices B-E were displayed in the identical format.

Further each choice was displayed in a random order to each participant. 
After completing their choices participants completed demographic information.

@@

Following \citeA{Read2005} and \citeA{Hardisty2015}, the dependent variable was the number of times the participant chose sooner reward (option 1 in Table~\ref{tab:stimuli}). 
As seen the Table~\ref{tab:study1results} (note I did not actually collect data for this study I just put sample data down for illustrative purposes) there is a strong effect of explicit intertemporal ratios. 
Nearly twice as many people choose the sooner option when temporal ratios are explicit.

\begin{table}[!ht]
	\caption{Proposed Percent of choices for the sooner option given different temporal ratios} 
	\label{tab:study1results}
\begin{tabular}{ p{3cm}||p{3cm}|p{3cm}  }
	\multicolumn{3}{|c|}{Temporal Frame} \\
	\hline
	 & Implict Ratio & Explicit Ratio\\
	\hline
	A (\%)	 & 30  & 60\\
	B (\%) 	 & 40  & 65   \\
	C (\%) 	 & 42  & 70\\
	D (\%)   & 37  & 68\\
	E (\%)   & 39  & 66\\
	Mean (\%) & 38  & 65.8 \\
	N  	 & 100  & 100\\
	\hline

\end{tabular}
\end{table}

Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis tests.
 Overall the effect of ratio was significant. 
 To confirm these results, I ran a robust logistic regression which allowed for correlated errors within participants.
 The dependent variable was choice (0 later option chosen, 1 sooner option chosen) and the independent variable was ratio frame. 
 This regression confirmed the results of the Kruskal-Wallis test, namely that participants were more impatient when temporal ratios are make explicit.

\section{oldstuff}

@@

%\begin{table}[!htbp] \centering 
%	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
%	\hline
%	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
%	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
%	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
%	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
%	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
%	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
%	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
%	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
%	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
%	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
%	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
%	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
%	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
%\end{tabular}
%\end{table}
